Multi-Index and Multi-Level Monte Carlo Evaluation of HJM Models

### 2016-03-27 tests/HJM/manuscript.tex
still compiling
and still working
now the friggin bjork paper works!
working with a reasonable text

\section{Introduction}
\label{section:Introduction}

%\subsection{Levy processes and finances}
\levy processes form a rich field within mathematical
finance. They allow modelling of asset prices with possibly discontinuous dynamics. An early and probably the best known model involving
a \levy process is
the \cite{Merton1976125} model, which generalises the \cite{black1973pricing} model. More recently, we have seen more complex models allowing
for more general dynamics of the asset price. Examples of such models include
the \cite{kou2002jump} model (see also \cite{Dotsis20073584}), the Normal Inverse Gaussian
model (\cite{barndorff1997normal, rydberg1997normal}),
the Variance Gamma model (\cite{madan1990variance, madan1998variance}),
and the Carr-Geman-Madan-Yor (CGMY) model
(\cite{cgmy_fine,carr2003stochastic}). For a good exposition on jump processes in finance we refer to \cite{conttankov}
(also see \cite{raible2000levy} and \cite{eberlein2001application}).

%\subsection{European option PIDE}
Prices of European options whose underlying asset is driven
by the \levy process are solutions to partial integro-differential
Equations (PIDEs) (\cite{nualart2001backward,briani_chioma_natalini,almendral2005numerical,kiessling2011diffusion})
that generalise
the Black-Scholes equation by incorporating a
non-local integral term to account for the discontinuities in
the asset price. This approach has also been extended to
cases where the option price features path dependence,
for instance in \cite{boyarchenko2002barrier} \cite{d2004penalty} and \cite{lord2008fast}.

%\subsection{Fourier methods}
The \levy-Khintchine formula provides an explicit 
representation of the characteristic function of a 
\levy process (cf, \cite{tankov2004financial}).
As a consequence, one can derive an exact expression for the 
Fourier transform of the solution of the relevant PIDE.
Using the inverse fast Fourier transform (iFFT) method, one
may efficiently compute the option price for a range of asset prices
simultaneously.
Furthermore, in the case of European call options, one may
use the duality property presented by \cite{dupire1997pricing}
and iFFT to efficiently compute option prices for a wide range of strike prices.

%\subsection{Need for an error bound and state of the art (Lee)}
Despite the popularity of Fourier methods for option pricing, few works
can be found on the error analysis and related parameter selection for these methods. 
A bound for the error not only provides an interval for the precise
value of the option, but also suggests a method to select the parameters of the numerical method.
An important
work in this direction is the one by \cite{lee2004option} in which several
payoff functions are considered for a rather general set of models, whose
characteristic function is assumed to be known.
\cite{fenglinetsky} presents the framework and theoretical approach
for the error analysis, and establishes polynomial convergence rates
for approximations of the option prices.
For a more
contemporary review on the error committed in various
FT-related methods we refer the reader to \cite{boyarchenko2011new},
that extends the classical \emph{flat} Fourier methods by deforming the integration countours
on the complex plane, studying discretely monitored barrier options
studied in \cite{de2014pricing}.

%\subsection{Our contribution and comparison with Lee}
In this work, we present a methodology for studying
and bounding  the error committed
when using FT methods to compute option prices.
We also provide a systematic way of choosing 
the parameters of the numerical method, in a way that minimises
the strict error bound, thus guaranteeing adherence to a pre-described
error tolerance. 
We focus on exponential \levy processes that may be of either
diffusive or pure jump. Our contribution is to derive a strict error bound for a Fourier
transform method when pricing options under risk-neutral 
\levy dynamics. We derive a simplified bound that separates the contributions of
the payoff and of the process in an easily processed and extensible
product form that is independent of the asymptotic behaviour of the
option price at extreme prices and at strike parameters.
We also provide a proof for the existence of optimal
parameters of the numerical computation that minimise the presented
error bound.
When comparing our work with Lee's work we find that Lee's work is more general than ours in that he studies a wider range
of processes, on the other hand, our results apply to a larger class of payoffs.
On test examples of practical relevance, we also find that the bound presented produces comparable or better results than the ones previously presented in the literature, with acceptable computational cost.

%\subsection{Organization of the paper}
The paper is organised in the following sections:
In Section \ref{section:Method}
we introduce the PIDE setting in the context of
risk-neutral asset pricing; we show the Fourier representation of the relevant
PIDE for asset pricing with \levy processes and use that
representation for derivative pricing.
In Section \ref{section:Bound}
we derive a representation
for the numerical error and divide it into
quadrature and cutoff contributions. We also describe the
methodology for choosing numerical
parameters to obtain minimal error bounds
for the FT method.
The derivation is supported by numerical examples
using relevant test cases with both diffusive and
pure-jump \levy processes in Section \ref{section:Numerics}.
Numerics are followed by conclusions in Section \ref{section:Conclusion}.

\section{Fourier method for option pricing}

\label{section:Method}
Consider an asset whose price at time $t$ is modelled 
by the stochastic process $S=(S_t)$ defined by $S_t=S_0 \expfs{X_t}$,
where $X=(X_t) \in \R$ is assumed to be
a \levy process whose jump measure $\nu$ satisfies
\begin{align} \label{eq:nuhyp}
\int_{\rz} \min\{y^2,1\} \nu \parent{dy} < \infty
\end{align}

Assuming the risk-neutral dynamic for $S_t$, the price at time $t=T-\tau$ of a European option
with payoff $G$ and maturity time $T$ is given by
\begin{align*}
\Pi(\tau,s)=\expfs{-r\tau} \expp{G\parent{S_T}|S_{T-\tau}=s}
\end{align*}
where $r$ is the short rate that we assume to be constant and $\tau\colon 0\leq\tau\leq T$ is the time to maturity. 
Extensions to non-constant deterministic short rates are straightforward.

The infinitesimal generator of a \levy
process $X$ is given by \citep[see][]{applebaum2004levy}
\begin{align}
\lx f(x)&\equiv 
\lim_{h\to 0} \frac{\expp{f(X_{t+h})|X_t=x}-f(x)}h \notag \\
&= \gamma f'(x)+\frac12\sigma^2f''(x)+\int_{\rz} \left(f(x+y)-f(x)-y1_{|y|\leq 1} f'(x)\right)\nu(\d y)
\end{align}
where $(\gamma,\sigma^2,\nu)$ is the characteristic triple of the \levy process.
The risk-neutral assumption on $(S_t)$ implies
\begin{equation} \label{eq:riskneutralexpint}
\int_{\absval{y}>1}\expfs{y} \nu(\d y)<\infty
\end{equation}
and fixes the drift term (see \cite{kiessling2011diffusion}) $\gamma$ of the \levy process to
% TODO: find more references to this
\begin{align} 
\label{drift_equation}
\gamma = r-\frac12 \sigma^2 - \int_{\rz} \parent{\expfs{y}-1-y1_{|y|\leq 1} }\nu(\d y)
\end{align}
Thus, the infinitesimal generator of $X$ may be written under the risk-neutral assumption as
\begin{align}
\lx f(x)&= \left(r-\frac{\sigma^2}2\right) f'(x)+\frac{\sigma^2}2 f''(x)+\int_{\rz} \left(f(x+y)-f(x)-(\expfs{y}-1)f'(x)\right)\nu(\d y)
\end{align}
Consider $g$ as the reward function in log prices (ie, defined by 
$g(x)=G(S_0 \expfs{x})$). Now, take $f$ to be defined as
\begin{align*}
f \parent{\tau,x} \equiv \expp{g \parent{X_{T}}|X_{T-\tau}=x}
\end{align*}
Then $f$ solves the following PIDE:
\begin{equation*}
\begin{cases}
\partial_{\tau}f(\tau,x)&=\lx f(\tau,x) \\
f(0,x)&=g(x), \qquad \qquad \parent{\tau,x}  \in [0,T] \times \R
\end{cases}
\end{equation*}
Observe that $f$ and $\Pi$ are related by
\begin{equation}
\Pi(\tau,S_0 \expfs{x})=\expfs{-r \tau}f(\tau,x)
\end{equation}
Consider a damped version of $f$ defined by $\fa(\tau,x)=\expfs{-\damp x}f(\tau,x)$; we see that $\partial_{\tau} \fa = \expfs{-\damp x} \lx f\parent{\tau,x} $.

There are different conventions for the Fourier transform. Here we consider the operator $\F$ such that 
\begin{align}
\F[f] \parent{\omega} \equiv \int_{\R}\expfs{i\omega x} f \parent{x} \d x
\label{eq:fourierDefinition}
\end{align}
defined for functions $f$ for which the previous integral is convergent. We also use $\ft{f} \parent{\omega}$ as a shorthand notation of $\F[f]\parent{\omega}$.
To recover the original function $f$, we define the inverse Fourier transform as 
$$\ift{f}(x)=\frac{1}{2\pi}\int_{\R} \expfs{-i\omega x}f(\omega) \d \omega$$
We have that $\ift{\ft{f}}(x)=f(x)$.

Applying $\F$ to $\fa$ we get $\ft{\fa}(\omega)=\ft{f} \parent{\omega + i \damp}$.
Observe also that the Fourier transform applied to $\lx f(\tau,x)$ gives $\charExp{-i\omega}\ft{f} \parent{\tau,\omega}$,
where $\charExp{\cdot}$ is the characteristic exponent of the process $X$, which satisfies $\expp{\expfs{zX_t}}=\expfs{t\charExp{z}}$.
The explicit expression for $\charExp{\cdot}$ is 
\begin{equation}
\label{characteristicDefinition}
\charExp{z}
=\parent{r-\frac{\sigma^2}2} z +\frac{\sigma^2}2 z^2 +\int_{\R} \left(\expfs{z y}-1-(\expfs{y}-1)z\right)\nu(\d y) \\[.5em]
\end{equation}
From the previous considerations it can be concluded that 
\begin{equation}
\partial_{\tau} \ft{\fa} = \charExp{\damp -i \omega} \ft{f} \parent{\omega - i \damp}
\end{equation}
Now $\ft{f}\parent{\omega - i \damp} = \ft{\fa} \parent{\omega}$ so $\ft{\fa}$ satisfies the following ODE
\begin{equation}
\begin{cases}
\frac{\partial_{\tau}\ft{\fa}\parent{\tau,\omega}}{\ft{\fa}(\tau,\omega)}&=\charExp{\damp -i \omega}\\[.5em]
\ft{\fa} \parent{0,\omega}&=\ft{\ga}\parent{\omega}
\end{cases}
\end{equation}

Solving the previous ODE explicitly, we obtain
\begin{equation}
\ft \fa \parent{\tau,\omega}
= \expfs{\tau \charExp{\damp-i \omega} } \ft{\ga} \parent{\omega}
\label{eq:theODE}
\end{equation}
Observe that the first factor in the right-hand side in the above 
equation is $\expp{\expfs{\parent{\damp-i \omega}X_{\tau}}}$, (ie, $\charFun{1}{-i\damp-\omega}$), where $\charFun{\tau}{\cdot}$ denotes the characteristic function of the
random variable $X_{\tau}$
\begin{align}
\charFun{\tau}{\omega} \equiv \expp{\tau \charExp{i \omega}} 
\end{align}

Now, to obtain the value function we employ
the inverse Fourier transformation, to obtain
\begin{equation}
\fa(\tau,x)=\ift{\ft{\fa}}(\tau,x)=\frac{1}{2\pi}\int_{\R} \expfs{-i\omega x} \ft{\fa}(\tau,\omega)d\omega
\label{eq:fadefinition}
\end{equation}
or
\begin{equation}
\fa(\tau,x)=\frac{1}{\pi}\int_{0}^{+\infty} \real{\expfs{-i\omega x} \ft{\fa}(\tau,\omega)}d\omega
\label{eq:fadefinition2}
\end{equation}

As it is typically not possible to compute the inverse Fourier transform analytically,
we approximate it by
discretising and truncating the integration domain
using trapezoidal quadrature \eqref{eq:fadefinition}.
Consider the following approximation:
\begin{align}
\label{eq:fdiscrete}
 \fadisc{\step,n} (\tau,x)&= \frac{\step}{2\pi} \sum_{k=-n}^{n-1} \expfs{-i{\parent{k+\frac12}\step} x} \ft{\fa}\parent{\tau,{\parent{k+\frac12}\step}}\\
 &=\frac{\step}{\pi} \sum_{k=0}^{n-1} \real{\expfs{-i{\parent{k+\frac12}\step} x} \ft{\fa}\parent{\tau,{\parent{k+\frac12}\step}}}
\end{align} 
Bounding and consequently minimising the error
in the approximation of $f(\tau,x)$ by $$\fdisc{\step,n}(\tau,x) \equiv \expfs{\damp x}\fadisc{\step,n}(\tau,x)$$
is the main focus of this paper and will be addressed in the following section.

\begin{remark}
Although we are mainly concerned with option pricing when 
the payoff function can be damped in order to guarantee
regularity in the $L^1$ sense, we note here that our main 
results are naturally extendable to include the Greeks of the option. 
Indeed, we have by \eqref{eq:theODE} that 
\begin{align}
 f \parent{t,x} = \frac{1}{2 \pi} \int_{\mathbb R} \expfs{\parent{\damp - i \omega}x} \ft \fa \parent{\tau, \omega} \d \omega
\end{align}
so the Delta and Gamma of the option equal
\begin{align}
\Delta \parent{t,x} &\equiv \frac{\partial f \parent{t,x}}{\partial x} = \frac{1}{2 \pi} \int_{\mathbb R} \parent{\damp-i \omega} \expfs{\parent{\damp - i \omega}x} \ft \fa \parent{\tau, \omega} \d \omega
\\
\Gamma \parent{t,x} &\equiv \frac{\partial^2 f \parent{t,x}}{\partial x^2} = \frac{1}{2 \pi} \int_{\mathbb R} \parent{\damp-i \omega}^2 \expfs{\parent{\damp - i \omega}x} \ft \fa \parent{\tau, \omega} \d \omega
%\\
%\Theta \parent{t,x}  &\equiv \frac{\partial f \parent{t,x}}{\partial t} = -\frac{1}{2 \pi} \int_{\mathbb R} \charExp{\damp-i \omega} \expfs{\parent{\damp - i \omega}x} \ft \fa \parent{\tau, \omega} d\omega
\end{align}
Because the expressions involve partial derivatives with respect to only $x$,
the results in this work are applicable for the computation of $\Delta$ and $\Gamma$
through a modification of the payoff function:
\begin{align}
\ft g_{\damp, \Delta} \parent{\omega} =& \ft g_{\damp} \parent{\omega} \parent{\damp - i \omega}
\\
\ft g_{\damp, \Gamma} \parent{\omega} =& \ft g_{\damp} \parent{\omega} \parent{\damp-i \omega}^2
\end{align}
When the Fourier space payoff function manifests
exponential decay, the introduction of a coefficient that is polynomial in $\omega$
does not change the regularity of $\ft g$ in a way that would significantly change the following
analysis. Last, we note that since we do our analysis for PIDEs on a mesh of
$x$'s, one may also compute the option values in one go and obtain the Greeks with
little additional effort using a finite difference approach for the derivatives.
\end{remark}

\subsection{Evaluation of the method for multiple values of $x$ simultaneously}

The Fast Fourier Transform (FFT) algorithm provides an efficient 
way of computing \eqref{eq:fdiscrete} for an equidistantly
spaced mesh of values for $x$ simultaneously.
Examples of works that consider this widely extended tool are \cite{lord2008fast,jackson2008fourier,hurd2010fourier} and \cite{schmelzle2010option}.
%In Subsection \ref{subsection:FFT} we analyse this alternative.

Similarly, one may define the Fourier frequency $\omega$
as the conjugate variable of some external parameter on which
the payoff depends. Especially, for the practically relevant case
of call options, we can denote the log-strike as $k$ and treat
$x$ as a constant and write:
\begin{align}
\kft f_{k,\damp} \parent{\omega} \equiv
\int_\R \expfs{\parent{\damp +i \omega }k } f_k \parent{x} \d k
\end{align}

Using this convention, the time dependence is given by
\begin{align}
\label{eq:kftDef}
\kft f_{k,\damp} \parent{\tau,x} = 
\frac{\expfs{\parent{i \omega + \damp + 1} x} \charFun{\tau}{\omega - i \parent{\damp +1} }}{\parent{i \omega + \damp} \parent{i \omega + \damp + 1}}
\end{align}
contrasted with the $x$-space solution
\begin{align}
\label{eq:xftDef}
\ft f_{k,\damp} \parent{\tau,x} = 
\frac{\expfs{\parent{i \omega - \damp + 1} k} \charFun{\tau}{\omega + i \damp }}{\parent{i \omega + \damp} \parent{i \omega + \damp + 1}}
\end{align}
We note that for call option payoff to be in $L^1$,
we demand that $\damp$ in \eqref{eq:kftDef} is positive.
Omitting the exponential factors that contain the $x$ and $k$
dependence in \eqref{eq:kftDef} and \eqref{eq:xftDef} respectively,
we have that one can arrive from \eqref{eq:kftDef} to \eqref{eq:xftDef}
using the mapping $\damp \mapsto -\damp-1 $. Thanks to this,
much of the analysis regarding the $x$-space transformation
generalises in a straightforward manner to the $k$-space transform.

\section{Error bound}
\label{section:Bound}

The aim of this section is to compute a bound of the error when approximating 
the option price $f(\tau,x)$ by $\fadisc{\step,n} (\tau,x)$,
defined in \eqref{eq:fdiscrete}.
Considering
\begin{equation} \label{eq:serie}
\fadisc{\step}(\tau,x)= \frac{\step}{2\pi} \sum_{k\in \Z} \expfs{-i{\parent{k+\frac12}\step} x} \ft{\fa}\parent{\tau,{\parent{k+\frac12}\step}}
\end{equation}
the total error $\Error$ can be split into a sum of two terms: the quadrature and truncation errors.
The former is the error from the approximation of the integral in \eqref{eq:fadefinition}
by the infinite sum in \eqref{eq:serie}, while the latter is due to the truncation of the infinite sum. 
Using triangle inequality, we have
\begin{align} \label{eq:errordef}
\Error &:=\absval{f(\tau,x)-\fdisc{\step,n}(\tau,x)}\leq \EQ + \EF
\end{align}
with
\begin{align*}
\EQ &= \expfs{\damp x}\absval{\fa(\tau,x)-\fadisc{\step}(\tau,x)}\\
\EF &= \expfs{\damp x} \absval{\fadisc{\step}(\tau,x)-\fadisc{\step,n}(\tau,x)}
\end{align*}
Observe that each $\Error$, $\EQ$ and $\EF$ depend on
three kinds of parameters:
\begin{itemize}
 \item Parameters underlying the model and payoff such as volatility and strike price. We call these \emph{physical parameters.}
 \item Parameters relating to the numerical scheme such as $\damp$ and $n$.
 \item \emph{Auxiliary parameters} that will be introduced in the process of deriving the error bound. These parameters
 do not enter the computation of the option price, 
 but they need to be chosen appropriately to have as tight a bound as possible. 
\end{itemize}

We start by analysing the quadrature error.
\subsection{Quadrature error}

Denote by $\strip{a}$, with $a>0$, the strip of width $2a$ around the real line:
$$\strip{a} \equiv \left\{z\in \C \colon \absval{\im{z}}<a \right\}$$
The following theorem presents conditions under which the quadrature error goes to zero at a spectral rate as $\step$ goes to zero.
Later in this section, we discuss simpler conditions to verify the hypotheses and analyse in more detail the case 
when the process $X$ is a diffusive process or there are ``enough small jumps."

\begin{theorem}\label{theo:quadrature} 
Assume that for $a > 0$:
\begin{itemize}
\item[H1.] the characteristic function of the random variable $X_1$ has an analytic extension to the set $$\strip{a}-\damp i \equiv \left\{z\in \C \colon \absval{\im{z}+\damp}<a\right\}$$
\item[H2.] the Fourier transform of $\ga(x)$ is analytic in the strip $\strip{a}$ and
\item[H3.] there exists a continuous function $\gamma \in L^1(\R)$ such that  $\absval{\ft{\fa}(\tau,\omega+i \beta)}<\gamma(\omega)$ for all $\omega\in \R$ and for all $\beta \in [-a,a]$
\end{itemize}
Then the quadrature error is bounded by
\begin{equation*}
\ErrorQuadrature \leq \expfs{\damp x} \frac{M_{\damp,a}(\tau,x) }{2\pi\parent{\expfs{2\pi a/\step}-1}}
\end{equation*}
where $M_{\damp,a}(\tau,x)$ is given by
\begin{align}\label{eq:defM}
M_{\damp,a}(\tau,x)&:=\sum_{\beta \in \{-a,a\}}\int_{\R}\absval{\expfs{-i\parent{\omega + i\beta} x}\ft{\fa}(\tau,\omega+i \beta)} \d \omega
\end{align}
$M_{\damp,a}(\tau,x)$ equals the Hardy norm (defined in \eqref{eq:hardynorm}) of the function $\omega \mapsto \expfs{-i\parent{\omega + i\beta} x}\ft{\fa}(\tau,\omega+i \beta)$, which
is finite.
\end{theorem}

The proof of Theorem \ref{theo:quadrature} is an application of Theorem 3.2.1 in \cite{stenger2012numerical},
whose relevant parts we include for ease of reading. Using the notation in \cite{stenger2012numerical}, $H^1_{\strip{a}}$ is the family of functions $w$ that are
 analytic in $\strip{a}$ and such that 
\begin{equation}\label{eq:hardynorm}
\nnorm{w}{H^1_{A_a}}:= \lim_{\varepsilon\to 0} \int_{\partial \strip{a}(\varepsilon)} \absval{w(z)}d\absval{z} < \infty
\end{equation}
 where 
 $$\strip{a}(\varepsilon) = \left\{z\in \C\colon \absval{\real{z}}<\frac1{\varepsilon},\ \absval{\im{z}} < a \parent{1-\varepsilon}\right\}$$
\begin{lemma}[Theorem 3.2.1 in \cite{stenger2012numerical}]\label{lem:stenger}
Let $w \in H^1_{A_a}$,
then define:
\begin{align}
I \parent{w} &= \int_\R w \parent{x} \d x \\
J \parent{w,h} &= h \ssum{j=-N}{N} w \parent{jh} \\
\zeta \parent{w,h} &= I \parent{w}- J \parent{w,h}
\end{align}
then
\begin{align}
\absval{\zeta \parent{w,h}} \leq \frac{e^{-\frac{\pi a}{h}} \nnorm{w}{H^1_{A_a}}}{2 \mathrm{sinh} \parent{\frac{\pi a}{h}}}
\end{align}
\end{lemma}

\begin{proof}[Proof of Theorem \ref{theo:quadrature}]
First observe that H1 and H2 imply that the function $w(z)=\expfs{-i x z + \step/2}\ft{\fa}(\tau,z+ \step/2)$ is analytic in $\strip{a}$. 
H3 allows us to use dominated convergence theorem to prove that $\nnorm{w}{H^1_{A_a}}$ is finite and coincides with $M_{\damp,a}(\tau,x)$. Applying Lemma \ref{lem:stenger} 
the proof is completed.
\end{proof}

Regarding the hypotheses of Theorem \ref{theo:quadrature}, the next propositions provide simpler conditions that imply H1 and H2 respectively.

\begin{proposition} \label{prop:analy1}
If $\alpha$, $a$ and $\nu$ are such that
\begin{equation}
\label{eq:alterH1}
\int_{y>1} \expfs{\parent{\damp+a}y}\nu(\d y) < \infty \quad \text{and}\quad \int_{y<-1} \expfs{\parent{\damp-a}y}\nu(\d y) < \infty
\end{equation}
then H1 in Theorem \ref{theo:quadrature} is fulfilled.

\begin{proof}
Denoting by $\charFun{1}{\cdot}$ the characteristic function of $X_1$, we want to prove that $z\mapsto \charFun{1}{z+\damp i}$ is analytic in $\strip{a}$. Considering that $\charFun{1}{z+\damp i}=\expfs{\charExp{iz-\damp}}$, the only non-trivial part of the proof is to verify that 
\begin{equation}
z\mapsto \int_{} p(z,y)\nu(\d y)
\end{equation}
is analytic in $\strip{a}$,
where $p\colon \strip{a} \times \R \to \C$ is given by
$$p(z,y)=\expfs{y \parent{i z-\damp}}-1-(\expfs{y}-1)\parent{i z-\damp}$$
To prove this fact, we show that we can apply the main result and the only theorem in \cite{mattner2001complex}, 
which, given a measure space $(\Omega,\mathcal{A},\mu)$ and an open subset $G\subseteq \C$, 
ensures the analyticity of $\int f(\cdot,\omega) d\mu(\omega)$, provided that $f\colon G\times \Omega \to \C$ satisfies: $f(z,\cdot)$ is $\mathcal{A}$-measurable for all $z\in G$; $f(\cdot,\omega)$ is holomorfic for all $\omega \in \Omega$; and $\int \absval{f(\cdot,\omega)} d\mu(\omega)$ is locally bounded.
In our case we consider the measure space to be $\R$ with the Borel $\sigma$-algebra and the Lebesgue measure, $G=\strip{a}$ and $f=p$. It is clear that 
$p(x,\cdot)$ is Borel measurable and $p(\cdot,y)$ is holomorphic. It remains to verify that  
\begin{equation*}
z \mapsto \int_{\R^*} \absval{p(z,y)}\nu(\d y)
\end{equation*} 
is locally bounded. 
To this end, we assume that $\real{z}<b$ (and, since $z\in \strip{a}$, $\im{z}<a$) and split the integration domain in $\absval{y}>1$ and ${0<\absval{y}\leq 1}$ to prove that both integrals are uniformly bounded.

Regarding the integral in ${\absval{y}> 1}$, we observe that
\begin{equation}
\absval{p(z,y)}\leq \expfs{y(\damp+\im{z})}+1+(\expfs{y}+1)(\damp+a+b)
\end{equation}
for $y<-1$ we have $\expfs{y(\damp+\im{z})}<\expfs{y(\damp-a)}$ while for $y>1$ we have $\expfs{y(\damp+\im{z})}<\expfs{y(\damp+a)}$. 
Using the previous bounds and the hypotheses together with \eqref{eq:nuhyp} and \eqref{eq:riskneutralexpint}, we obtain the needed bound.

For the integral in ${0<\absval{y}\leq 1}$, observe that, denoting $f(z,y)=\absval{p(z,y)}$, we have, $f(z,0)=0$ for every $z$, $\partial_y f(z,0)=0$ for every $z$, and $\absval{\partial_{yy}f(z,y)}< c$ for $z\in \strip{a}, \real{z}<b, \absval{y}<1$. From these observations we get that the McLaurin polynomial of degree one of $y\mapsto f(z,y)$ is null for every $z$, and we can bound $f(z,y)$ by the remainder term, which, in our region of interest, is bounded by $\frac{c}{2} y^2$, obtaining
\begin{equation}
\int_{0<\absval{y}\leq 1} \absval{p(z,y)}\nu(\d y)\leq \frac{c}2 \int_{0<\absval{y}\leq 1}y^2\nu(\d y)
\end{equation}
which is finite by hypothesis on $\nu$, which finishes the proof.
\end{proof}
\end{proposition}

\begin{proposition}
If for all $b<a$, the function $x\mapsto \expfs{b\absval{x}}{\ga(x)}$ is in $L^2(\R)$ then H2 in Theorem \ref{theo:quadrature} is fulfilled.
\end{proposition}
\begin{proof}
The proof is a direct application of Theorem IX.13 in \cite{reedsimon}
\end{proof}
We now turn our attention to a more restricted class of Levy processes. 
Namely, processes such that either $\sigma^2>0$ or there exists $\lambda \in (0,2)$ such that $\cl$ defined in \eqref{eq:PureJumpC} is strictly positive. 
For this class of processes, we can state our main result explicitly in terms of the characteristic triplet.

Given $\lambda \in (0,2)$, define $\cl$ as
\begin{equation}
\label{eq:PureJumpC}
\cl  = 
\textrm{inf}_{\kappa > 1} \left\{\kappa^{\lambda} 
 \int_{0<|y| < \frac{1}{\kappa}} y^2 \nu(\d y)\right\}
\end{equation}
Observe that $\cl\geq 0$ and, by our assumptions on the jump measure $\nu$, $\cl$ is finite. Furthermore, if $\lambda \in (0,2)$ is such
that
\begin{equation}
\label{eq:pureJumpAssumption}
\liminf_{\epsilon \downarrow 0} \frac{1}{\epsilon^{\lambda}}
 \int_{0<|y| < \epsilon} y^2 \nu(\d y) > 0
\end{equation}
then $\cl>0.$ To see this, observe that \eqref{eq:pureJumpAssumption} implies the existence of $\epsilon_0$ such that 
\begin{equation*}
\inf_{\epsilon \leq \epsilon_0}\left\{ \frac{1}{\epsilon^{\lambda}}
 \int_{0<|y| < \epsilon} y^2 \nu(\d y)\right\} > 0
\end{equation*}
If $\epsilon_0<1$ observe that
\begin{equation*}
\inf_{\epsilon_0 \leq \epsilon \leq 1}\left\{ \frac{1}{\epsilon^{\lambda}}
 \int_{0<|y| < \epsilon} y^2 \nu(\d y)\right\} \geq  \int_{0<|y| < \epsilon_0} y^2 \nu(\d y) > 0
\end{equation*}
where for the first inequality it was taken into account that $\frac{1}{\epsilon^{\lambda}}\geq 1$ and that the integral is increasing with $\epsilon$.
Combining the two previous infima and considering $\absval{\kappa}=\frac{1}{\epsilon}$ we get that $\cl>0.$

Furthermore, we note that for a \levy model with finite jump
intensity, such as the Black-Scholes and Merton models that satisfy the first of our
assumption, $\cl=0$ for all $\lambda \in (0,2)$.

\begin{theorem} \label{teo:boundM}
Assume that: $\damp$ and $a$ are such that \eqref{eq:alterH1} holds; $\ft{\ga} \in \linf{\strip{a}}$; and either $\sigma^2>0$ or $\cl>0$ for some $\lambda\in (0,2)$.
Then the quadrature error is bounded by
\begin{equation*}
\ErrorQuadrature \leq \expfs{\damp x} \frac{\tilde{M}_{\damp,a}(\tau,x) }{2\pi\parent{\expfs{2\pi a/\step}-1}}
\end{equation*}
where 
\begin{equation}\label{eq:Mbound}
\tilde{M}_{\damp,a} \parent{\tau,x} 
=
\ssum{c\in \sset{-1,1}}{} 
\expfs{cax}
\expfs{\tau \Psi \parent{ca}} \absval{\ft{\ga} \parent{ca}}
\int_{\R}  \expfs{-\tau \left( \frac{\sigma^2}{2}\omega^2 + \frac{|\omega|^{2 - \lambda}}{4} \cl\indicator_{\absval{\omega}>1}\right) } \d \omega
\end{equation}
Furthermore, if $\sigma^2>0$ we have
\begin{equation}
\label{eq:Mboundsigpositive}
\tilde{M}_{\damp,a} \parent{\tau,x}  \leq \frac{\sqrt{2\pi}}{\sigma\sqrt{\tau}}\ssum{c\in \sset{-1,1}}{} 
\expfs{cax}
\expfs{\tau \Psi \parent{ca}} \absval{\ft{\ga} \parent{ca}}
\end{equation}

>>>>

We are interested in analysing the Forward Euler Finite Element evaluation of forward curve models of HJM models. The main approach was represented with a convergence analysis in \cite{bjork2013monte}, whose notation we largely follow.

We are focused on the evolution of the forward curve $f \parent{t , \tau}$ for $t \in [0,t_{max}]$, $\tau \in [0,\tau_{max}]$ with $\tau_{max}>t_{max}$. In order to define a mesh for the numerical solution of the problem, let us define $\bar \ell \in \mathbb{N}^3$.

Let us first discretise the time interval as
\begin{align*}
\Delta t_{\bar \ell} &= 2^{-\ell_1} t_{max},
\\
t_{\bar \ell, n} &= n \Delta t_{\bar \ell}, ~~~ n \in [0,1,2,\cdots, 2^{\ell_1}].
\end{align*}
Since there is an additional requirement that $t_{max}$ be in
the mesh of maturities, one point in the mesh in $\tau \in [0,\tau_{max}]$. We let $\ell_1$ and $\ell_2$ define the number of discretisation points to the right and to the left of the fixed
point, respectively.
\begin{align*}
\Delta \tau_{\bar \ell,1} &= 2^{-\ell_3} t_{max},  \\
\Delta \tau_{\bar \ell,2} &= 2^{-\ell_2} \parent{\tau_{max}-t_{max}}, \\
\tau_{\bar \ell, n} &= n \Delta \tau t_{\bar \ell,1}, ~~~ n \in [0,1,2,\cdots, 2^{\ell_3}] \\
\tau_{\bar \ell, n} &= n \Delta \tau t_{\bar \ell,1}, ~~~ n \in [2^{\ell_3},2^{\ell_3}+1,\cdots, 2^{\ell_3}+2^{\ell 2}].
\end{align*}

Using this notation, we have the following rates of convergence, following the Multi-Index notation:
\begin{align}
\bar \beta &= [2,2,2]^T \\
\bar \gamma &= [1,1,1]^T .
\end{align} 
Choosing the Simpson method for integration, we have the order of quadrature $p_Q=3$, which brings the quadrature error to 
\begin{align}
\expp{\absval{\mathcal G\parent{\bar{\bar g}} - \mathcal G \parent{g}}^{2 \kappa}} \leq C \parent{\Delta \tau_{\bar \ell}}^{2 \kappa p_Q} = 
\tilde C 2^{-6 \kappa \mathrm{min} \parent{\ell_3,\ell_2}}.
\end{align}
On the other hand, the error of the simpson method is the sum of errors for $\tau < t_{max}$ and $\tau \leq t_{max}$. The error for these two domains are $\bigo{2^{-2 \ell_2}}$ and $\bigo {2^{-2 \ell_3}}$, respectively. Combining together, we get the weak and the strong rate
for the Simpson quadrature:
\begin{align}
\expp{\absval{\mathcal G\parent{\bar{\bar g}} - \mathcal G \parent{g}}^{2}} \leq&  \tilde C_Q 2^{-6 \mathrm{min} \parent{\ell_3,\ell_2}}
\\
\expp{\absval{\mathcal G\parent{\bar{\bar g}} - \mathcal G \parent{g}}} \leq& C_Q 2^{-2 \mathrm{min} \parent{\ell_2,\ell_3}}.
\end{align}

\begin{align}
\expp{\absval{\mathcal G\parent{\bar{\bar g}} - \mathcal G \parent{g}}^{2 \kappa}}
\end{align}
using Theorem 3.3 of \cite{bjork2013monte}.

On the other hand, theorem 3.2 and equation (3.54) states that
\begin{align}
\expp{ \absval{ \mathcal G \parent{g} - \mathcal G \parent{\bar{\bar g}} }^{2 \kappa}  } ^{\frac{1}{2 \kappa}} \leq C_{5}^{CV} \parent{ \parent{\Delta t}^{\frac{1}{2}} + \Delta \tau }.
\end{align}

The weak rate for the discretisation error can be bounded by eq. (4.6):
\begin{align}
\expp{ \absval{ \mathcal G \parent{g} - \mathcal G \parent{\bar{\bar g}} }^{}  }
= 
 \bigo{\parent{\Delta t}^2 + \parent{\Delta \tau}^2}
 = \bigo{2^{-2 \ell_1} + 2^{-2 \mathrm {min} \parent{\ell_2,\ell_3} }}
\end{align}

Due to the fact that the discretisation errors relating to $\tau$
depend on the minimum of $\ell_2$ and $\ell_3$, it seems natural
to assume $\ell_3=\ell_2$. This gives us a two-dimensional
Multi-Index Monte Carlo formulation with the following effective parameters
\begin{align}
\bar \beta &= [2,2]^T ,\\
\bar \gamma &= [1,1]^T , \\
\bar s &= [2^{-1},1], \\
\bar w & = [2,2].
\end{align}

On the other hand, one may set $\ell_1=\ell_2=\ell_3=\ell$
and arrive at a one-dimensional MLMC formulation with a weak rate of $2$, strong rate of $1$, and computational cost per realisation of $W_\ell=2^{2\ell}$. In the sequel, we will present the results of these two variations for comparison.

\section{Multi-Index Monte Carlo}

\section{Multi-Level Monte Carlo}
